---
title: "Exponential Distribution"
author: "Abhijith Asok, Chris Hilger, Liam Loscalzo, Katherine Wang"
output: beamer_presentation
theme: "Boadilla"
##colortheme: "structure"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(parallel)
```

# Point Estimation: Methods

We compared the following point estimators:

- Maximum Likelihood Estimator
- Unbiased correction for the MLE
- Second Method of Moment Estimator

##add criteria for comparing estimators!

# Maximum Likelihood Estimator
```{r}
##code here

```

# Unbiased correction for the MLE
```{r}
##code here
```

# Second Method of Moment Estimator
```{r}
##code here
```

# Confidence Intervals: Methods

We compared the following confidence intervals:

- Wald-based Confidence Interval
- Gamma-based Confidence Interval
- Score-based Confidence Interval
- Bootstrap Confidence Interval

##add criteria for comparing CIs!!

# Wald Confidence Interval

```{r, echo = TRUE}
wald_ci <- function(N, rate, alpha = 0.05){
  x <- rexp(N, rate = rate)
  x_bar <- mean(x)
  se <- sd(x)/sqrt(N)
  ci <- x_bar + c(-1, 1)*qnorm(1 - (alpha / 2))
  return(ci)
}
```

# Gamma Confidence Interval

If $X_1, X_2,...,X_n \stackrel{\text{iid}}{\sim} \text{Exponential}(\lambda)$, then $\sum_i^n x_i \sim \text{Gamma}(n, \lambda) \implies \lambda \bar{x} \sim \text{Gamma}(n, n)$. Let $g_{y}$ be the $y$th percentile of this distribution. Then we can say:

$$
1 - \alpha = P(g_{\alpha/2} \le \lambda \bar{x} \le g_{1-\alpha/2}) = P(g_{\alpha/2}/\bar{x} \le \lambda \bar{x} \le g_{1-\alpha/2}/\bar{x})
$$

And therefore, a $(1 - \alpha)\%$ confidence interval for $\lambda$ is 
$$
(g_{\alpha / 2}/\bar{x},\, g_{1 - \alpha / 2}/\bar{x})
$$


```{r, echo = TRUE}
gamma_ci <- function(N, rate, alpha = 0.05){
  # Inspiration: https://math.stackexchange.com/questions/1288139/calculate-the-confidence-interval-of-parameter-of-exponential-distribution
  x <- rexp(N, rate = rate)
  x_bar <- mean(x)
  ci_rate <- qgamma(c(alpha / 2, 1 - (alpha / 2)), N, N) / x_bar
  ci_mean <- 1 / ci_rate
  return(c(ci_mean[2], ci_mean[1]))
}
```

# Score Confidence Interval
```{r, echo = TRUE}
score_ci <- function(N, rate, alpha = 0.05) {
  x <- rexp(N, rate = rate)
  xbar <- mean(x)
  ci_lambda <- (1/xbar)*(1 + c(-1, 1) * qnorm(1 - alpha / 2) / sqrt(n))
  ci_mean <- 1 / ci_lambda
  return(c(ci_mean[2], ci_mean[1]))
}
```

# Bootstrap Confidence Interval

```{r, echo = TRUE}
bootstrap_ci <- function(N, rate, alpha = 0.05){
  # Function to calculate bootstrap CI
  x <- rexp(N, rate = rate)
  x_bar <- mean(x)
  # Number of bootstrap samples
  nb <- 1000
  # Take boostrap samples
  bootstrap_samples <- sample(x, N * nb, replace = TRUE) %>%
    matrix(nrow = N, ncol = nb)
  # Get means of columns 
  means <- colMeans(bootstrap_samples)
  # Get deltas (x* - x)
  deltas <- means - x_bar
  deltas <- sort(deltas)
  # Calculate CIs
  ci <- x_bar - quantile(deltas, probs = c(alpha / 2, 1 - (alpha / 2)))
  return(c(ci[2], ci[1]))
}
```

# Other

```{r, echo = FALSE}
library(dplyr)
# Other functions
coverage_probability <- function(N, rate, ci_fun, alpha = 0.05, B = 1000){
  # Match input function to actual function
  fun <- tryCatch(match.fun(ci_fun), 
                  error = function(e) print(paste0("ci_fun: '", ci_fun, "' does not exist")))
  # Expected value
  exp_val <- 1 / rate
  # Calculate B confidence intervals and put in dataframe
  conf_ints <- replicate(B, fun(N, rate)) %>%
    t() %>% data.frame()
  names(conf_ints) <- c("lower", "upper")
  # Calculate hit rates (expected value inside confidence interval)
  conf_ints <- mutate(conf_ints, hit = (exp_val >= lower & exp_val <= upper))
  coverage <- conf_ints$hit %>% mean()
  return(coverage)
}

# Work
set.seed(123)

lambda <- 1 / 12
beta <- 1 / lambda
n <- 60

# Coverage probability plotting
ns <- seq(1, 500)

###################################################
# I have commented out all of these so that we don't 
# accidentally re run if we don't have to. I've saved 
# the data for coverage probabilities so we can read it 
# in without re-running if not necessary.
###################################################
# Caution: bootstrap takes for-ev-er
# bootstrap_cov <- mclapply(ns, coverage_probability, rate = lambda, ci_fun = "bootstrap_ci") %>% unlist()
# gamma_cov <- mclapply(ns, coverage_probability, rate = lambda, ci_fun = "gamma_ci") %>% unlist()
# wald_cov <- mclapply(ns, coverage_probability, rate = lambda, ci_fun = "wald_ci") %>% unlist()
# score_cov <- mclapply(ns, coverage_probability, rate = lambda, ci_fun = "score_ci") %>% unlist()

#coverage_df <- data.frame(n = ns, bs = bootstrap_cov, gamma = gamma_cov, wald = wald_cov, score = score_cov)
#write_csv(coverage_df, "./data/coverage_probabilities.csv")
###################################################
# If need to read:
coverage_df <- read_csv("./data/coverage_probabilities.csv")

ggplot(coverage_df, aes(x = n, y = bs)) +
  geom_line(se = FALSE) +
  geom_line(aes(x = n, y = gamma)) +
  geom_line(aes(x = n, y = wald)) +
  geom_line(aes(x = n, y = score))
```

# Summary of Findings and Recommendations

```{r cars, echo = TRUE}
summary(cars)
```


